Find the E-value distribution of HMM Hits for HMM families in HMMER-DB.
=======================================================================
```{r message = FALSE}
# Imports:
library(ggplot2)
library(scales)
library(RSQLite)
```

#### Custom function for reverse logarithmic scale in the plot below:
```{r}
# Functions One: Creates a scale function for a reverse logarithmic scale.
reverselog_trans = function(base = exp(1)) 
{
  trans = function(x) -log(x, base)
  inv = function(x) base^(-x)
  trans_new(paste0("reverselog-", format(base)), trans, inv, 
            log_breaks(base = base), 
            domain = c(1e-100, Inf))
}
```

#### We can query the HMMER-DB database using a SQL query within R. Our first step is to setup a database connection:
```{r message = FALSE}
# Setting up database connection:
sqlite = dbDriver("SQLite")
HMMDB  = dbConnect(sqlite, "/Users/lee/Data/SteriodHMMs/HMMDBV2.sqlite") # Location of HMMER-DB Sqlite database.
```

#### The code below queries the database and loads the results of the query into a R dataframe. Please note the substr() function in the SQL. It is what allows us to select what phylogenetic level we want to explore.
```{r}
# Executes SQL query and loads results directly into a dataframe.
data = dbGetQuery(HMMDB, "SELECT
                            HMM_Hits.HMM_E_Value,
                            HMM_Hits.HMM_Model,
                          	HMM_Data.HMM_Family
                          FROM
	                          HMM_Hits,
	                          HMM_Data
                          WHERE
	                          HMM_Hits.HMM_Model = HMM_Data.HMM_Model
                            AND HMM_Hits.HMM_Coverage > 0.50") # Coverage filter 
```

#### We now check if the SQL query is generating the dataframe we want:
```{r}
head(data) # Note: The columns may wrap to the next line.
```

#### We can then use ggplot to plot the E-value distribution of the HMM Hits for HMM families in the database:
```{r fig.width = 10, fig.height = 25}
# Plots Data as a bar graph.
plotObj = ggplot(data, aes(x = HMM_E_Value, y = HMM_Model, colour = factor(HMM_Family)))
plotObj + geom_point(alpha = 2/5) + scale_x_continuous(trans = reverselog_trans(10)) + 
          facet_grid(HMM_Family ~ ., scales = "free") +
          theme(legend.position = "none") +
          ggtitle("E-Value distribution of all hits for all Hidden Markov Models.") + 
          ylab("HMM Model") + xlab("HMM hit E-Value distribution per HMM")
```