Count the number of HMM hits per HMM family per taxa in HMMER-DB.
=================================================================
```{r message = FALSE}
# Imports:
library(ggplot2)
library(RSQLite)
```

#### We can query the HMMER-DB database using a SQL query within R. Our first step is to setup a database connection:
```{r message = FALSE}
# Setting up database connection:
sqlite = dbDriver("SQLite")
HMMDB  = dbConnect(sqlite, "/Users/lee/Data/SteriodHMMs/HMMDBV2.sqlite") # Location of HMMER-DB Sqlite database.
```

#### The code below queries the database and loads the results of the query into a R dataframe. Please note the substr() function in the SQL. It is what allows us to select what phylogenetic level we want to explore.
```{r}
# Executes SQL query and loads results directly into a dataframe.
data = dbGetQuery(HMMDB, "/* SQL Query: Counts the number of distinct HMM families per organism source. */
                          SELECT
                              HMM_Data.HMM_Family,
                              count(HMM_Data.HMM_Family) AS Count,
	                            substr(Organisms.Organism_Phylogeny,0,18) AS Phylum /* Feather the 18 to change the Phylogeny; Less:HigherTaxa <-> Greater:LowerTaxa */
                          FROM
	                            HMM_Data,
	                            HMM_Hits,
	                            Organisms,
	                            Proteins
                          WHERE
	                            HMM_Data.HMM_Model = HMM_Hits.HMM_Model
                          AND HMM_Hits.Protein_Accession = Proteins.Protein_Accession
                          AND Proteins.Organism_Accession = Organisms.Organism_Accession
                          AND Organisms.Organism_Phylogeny LIKE 'Bacteria%' /* Use this to filter by specific Phylogeny */
                          GROUP BY
	                            HMM_Data.HMM_Family,
	                            Phylum
                          ORDER BY
	                            Phylum")

```

#### We now check if the SQL query is generating the dataframe we want:
```{r}
head(data) # Note: The columns may wrap to the next line.
```

#### We can then use ggplot to plot the number of hits for each HMM family within each phylum in the database:
```{r fig.width = 20, fig.height = 30}
# Plots data as a point graph.
plotObj = ggplot(data, aes(x = HMM_Family, y = Count, colour = factor(Phylum), size = 3))
plotObj + geom_point() + facet_grid(Phylum ~ ., scales = "free") +
          theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) + 
          ggtitle("Number of hits for HMM families in each phylum in the database.") + 
          ylab("Hit Count") + xlab("HMM Family")
```