Find the number of organisms with hits for particular HMM families in HMMER-DB.
===============================================================================
```{r message = FALSE}
#Imports:
library(ggplot2)
library(RSQLite)
```

#### We can query the HMMER-DB database using a SQL query within R. Our first step is to setup a database connection:
```{r message = FALSE}
# Setting up database connection:
sqlite = dbDriver("SQLite")
HMMDB  = dbConnect(sqlite, "/Users/lee/Dropbox/RandD/Repositories/HMMER-DB/TestDBs/TestDB.sqlite") # Location of HMMER-DB Sqlite database.
```

#### The code below queries the database and loads the results of the query into a R dataframe. Please note the substr() function in the SQL. It is what allows us to select what phylogenetic level we want to explore.
```{r}
# Executes SQL query and loads results directly into a dataframe.
data = dbGetQuery(HMMDB, "SELECT
                              HMM_Data.HMM_Family,
                              count(DISTINCT Organisms.Source) AS Organism_Count,
	                            substr(Organisms.Organism_Phylogeny,0,8) AS Kingdom  /* Feather the 8 to change the Phylogeny: Less:HigherTaxa <-> Greater:LowerTaxa */
                          FROM
	                            HMM_Data,
	                            HMM_Hits,
	                            Organisms,
	                            Proteins
                          WHERE
	                            HMM_Data.HMM_Model = HMM_Hits.HMM_Model
                          AND HMM_Hits.Protein_Accession = Proteins.Protein_Accession
                          AND Proteins.Organism_Accession = Organisms.Organism_Accession
                          GROUP BY
	                            Kingdom,
	                            HMM_Data.HMM_Family
                          ORDER BY
	                            HMM_Data.HMM_Family,
	                            Organism_Count DESC")
```

#### We now check if the SQL query is generating the dataframe we want:
```{r}
head(data) # Note: The columns may wrap to the next line.
```

#### We can then use ggplot to plot the number of organisms with hits for particular HMM families in the database:
```{r fig.width = 15, fig.height = 6}
# Plots Data as a bar graph.
plotObj = ggplot(data, aes(x = HMM_Family, y = Organism_Count, fill = Kingdom))
plotObj + geom_bar(stat = "identity", position = "identity") + coord_flip() +
          ggtitle("Number of organisms with hits for particular HMM families for each Kingdom") + 
          ylab("Number of organisms with hits.") + xlab("HMM Family")
```