Query for the relative positions of multiple Hidden Markov model hits across multiple genomes in HMMER-DB.
==========================================================================================================
```{r message = FALSE}
# Imports:
library(ggplot2)
library(RSQLite)
```

#### We can query the HMMER-DB database using a SQL query within R. Our first step is to setup a database connection:
```{r}
# Setting up database connection:
sqlite = dbDriver("SQLite")
HMMDB  = dbConnect(sqlite, "/Users/lee/Data/SteriodHMMs/OldDBs/HMMDBV4.sqlite") # Location of HMMER-DB Sqlite database.
```
#### The code below queries the database and loads the results of the query into a R dataframe (Note: This is a rather complex query with multiple filters). 
```{r}
# Executes SQL query and loads results directly into a dataframe. 
data = dbGetQuery(HMMDB, "/* SQL Outer Query: Wrapper for subquery one that calculates relative protein center from subquery one's relative protein start and stop.*/
SELECT
  subQuery.HMM_Family,
	subQuery.Organism_Description,
	(( subQuery.Protein_Relative_Start + subQuery.Protein_Relative_End ) / 2 ) AS Protein_Relative_Center
FROM
	(
		SELECT DISTINCT
			HMM_Data.HMM_Family,
			Organisms.Organism_Description,
			( CAST (Proteins.'End' AS float) / CAST ( Organisms.Sequence_Length AS float ) * 100 ) AS Protein_Relative_End, 
			( CAST (Proteins.Start AS float) / CAST ( Organisms.Sequence_Length AS float ) * 100 ) AS Protein_Relative_Start
		FROM
			HMM_Data,
			HMM_Hits,
			Organisms,
			Proteins
		WHERE
			HMM_Data.HMM_Model = HMM_Hits.HMM_Model
		AND HMM_Hits.Protein_Accession = Proteins.Protein_Accession
		AND Proteins.Organism_Accession = Organisms.Organism_Accession
		AND Organisms.Organism_Accession IN (
			/* SQL Inner Query 1: Selects organisms from the subquery with a HMM_Coverage of at least 60% for key steriod degrading enzymes.*/
			SELECT
				Organisms.Organism_Accession
			FROM
				HMM_Data,
				HMM_Hits,
				Organisms,
				Proteins
			WHERE
				HMM_Data.HMM_Model = HMM_Hits.HMM_Model
			AND HMM_Hits.Protein_Accession = Proteins.Protein_Accession
			AND Proteins.Organism_Accession = Organisms.Organism_Accession
			AND HMM_Data.HMM_Family IN (
				'hsaA',
				'hsaC',
				'hsaD',
				'hsaE',
				'hsaF',
				'hsaG',
				'hsd4B',
				'KshA',
				'kstD'
			)
			AND HMM_Hits.HMM_Coverage >= 0.8
			AND Organisms.Organism_Accession IN (
				SELECT DISTINCT
					Organisms.Organism_Accession
				FROM
					HMM_Hits,
					Organisms,
					Proteins
				WHERE
					HMM_Hits.Protein_Accession = Proteins.Protein_Accession
				AND Organisms.Organism_Accession = Proteins.Organism_Accession
				AND HMM_Hits.HMM_Model LIKE 'hsaC%' /* Selects organisms with a hit for HsaC*/
				AND HMM_Hits.HMM_Coverage >= 0.8 /* Select only hsaC hits with over 80% HMM coverage */
				AND Organisms.Organism_Accession IN (
					SELECT DISTINCT
						Organisms.Organism_Accession
					FROM
						HMM_Hits,
						Organisms,
						Proteins
					WHERE
						HMM_Hits.Protein_Accession = Proteins.Protein_Accession
					AND Organisms.Organism_Accession = Proteins.Organism_Accession
					AND HMM_Hits.HMM_Model LIKE 'KshA%' /* Selects organisms with a hit for KshA*/
					AND HMM_Hits.HMM_Coverage >= 0.8
					AND Organisms.Organism_Description NOT LIKE 'Myco%'
				)
			)
			GROUP BY
				Organisms.Source
			HAVING
				count(DISTINCT HMM_Data.HMM_Family) >= 5.4
			ORDER BY
				Organisms.Organism_Phylogeny ASC
		)
		AND HMM_Data.HMM_Family IN (
			'hsaA',
			'hsaC',
			'hsaD',
			'hsaE',
			'hsaF',
			'hsaG',
			'hsd4B',
			'KshA',
			'kstD'
		)
	) AS subQuery
ORDER BY
	subQuery.Organism_Description ASC,
	(( subQuery.Protein_Relative_Start + subQuery.Protein_Relative_End ) / 2 )")
```

#### We now check if the SQL query is generating the dataframe we want:
```{r}
head(data) # Note: The final column may wrap to the next line.
```

#### We can then use ggplot to plot the relative positions of HMM hits in the genomes of the organisms in the database:
```{r fig.width = 15, fig.height = 10}
plotObj = ggplot(data, aes(x = Protein_Relative_Center, y = Organism_Description, color = factor(HMM_Family)))
plotObj + geom_point(alpha = 3/4) + # Slight alpha so one can visualize overlaping points better.
          ggtitle("Relative positions of proteins with HMM hits for non-Mycobacterium bacteria in the database.") + 
          xlab("Protein's relative postion on geneome.") + ylab("Organism") + labs(colour = "Type of HMM hit:") 
```

*Please note that most bacterial genomes are circular and the genome may start at an arbitrary point.*

#### We can also use ggplot to plot the relative density of HMM hits in the genomes of the organisms in the database:
```{r fig.width = 15, fig.height = 10}
plotObj = ggplot(data, aes(x = Protein_Relative_Center, y = Organism_Description, fill = 'black'))
plotObj + geom_point(alpha = 1/5) + # Proteins are all the same color and we use alpha to highlight areas with a greater number of hit proteins.
          ggtitle("Relative density of proteins with HMM hits for non-Mycobacterium bacteria in the database.") + 
          xlab("Protein's relative postion on geneome.") + ylab("Organism") + theme(legend.position = "none")
```

*Please note that most bacterial genomes are circular and the genome may start at an arbitrary point.*

#### Lets say we want to filter the hits down further to just the key enzymes such as HsaC and KshA. We can use R to make a new filtred dataframe.
```{r}
dataTwo = data[data$HMM_Family %in% c("KshA", "hsaC"), ] # Select rows where the column HMM_Family is equal to KshA or hsaC.
```
#### We now check if the new dataframe is what we want:
```{r}
head(dataTwo) # Note: The final column may wrap to the next line.
```

#### We again use ggplot to plot the relative positions of HMM hits in the genomes of the organisms in the database but now only for HsaC and KshA:
```{r fig.width = 15, fig.height = 10}
plotObj = ggplot(dataTwo, aes(x = Protein_Relative_Center, y = Organism_Description, color = factor(HMM_Family)))
plotObj + geom_point(alpha = 3/4) + # Slight alpha so one can visualize overlaping points better.
          ggtitle("Relative positions of HsaC and KshA HMM hits for non-Mycobacterium bacteria in the database.") + 
          xlab("Protein's relative postion on geneome.") + ylab("Organism") + labs(colour = "Type of HMM hit:") 
```
